---
title: "Stat380 Mini-Project 2"
author: "Ben Rishel and Alven Huang"
date: "October 11, 2025"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Discussion of Group Member Contributions

The work was divided evenly between us. We contributed via a Github repository. After the work was completed, each of us reviewed the full document before submission to ensure consistency. Ben completed Task 1 and Tasks 4 and 5. Alven completed the Tasks 2 and 3.

## Front Matter

```{r}
rm(list = ls())
library(tidyverse)
library(readxl)
library(kableExtra)
library(ggplot2)
library(lubridate)

cod <- read_excel("CODGames2_mp.xlsx", 
  col_types = c("text", "text", "text", 
  "text", "date", "text", "text", "numeric", 
  "numeric", "numeric", "numeric", 
  "numeric", "text", "text", "text", 
  "text", "numeric", "numeric", "numeric", 
  "numeric", "numeric", "numeric", 
  "numeric", "numeric", "numeric", 
  "numeric", "numeric"))

```

## Task 1 - Why you should avoid using na.omit()

Let's investigate why it is a bad idea to use na.omit() to remove observations with missing data for this dataset. First, let's look at how many NA values we have per column through the use of an anonymous function.

```{r}
sapply(cod, function(x) sum(is.na(x)))
```

If we used na.omit() on this dataset, we would remove any row that has at least one NA value. In this instance, that would result in the entire dataset being removed because there are 211 observations in the dataset and also 211 NA values for diffuses, plants, detonates, and deposits. A better approach in this instance would be to remove just the columns with so many NA values. Now, let's make confirm that there are NA values in every row.

```{r}
apply(cod, 1, function(x) sum(is.na(x)))
```

When counting the NA values by row, you can clearly see that there are many NA values per each row. This supports the previously mentioned point that every single row would be omitted with na.omit().

## Task 2 - 3 More Univariate Explorations

We previously explored Eliminations, PrimaryWeapon, Damage, Deaths, and TotalXP in Mini-Project 1.
