---
title: "Stat380 Mini-Project 3"
author: "Ben Rishel and Alven Huang"
date: "2025-12-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Discussion of Group Member Contributions

The work was divided evenly between us. We contributed via a Github repository. After the work was completed, each of us reviewed the full document before submission to ensure consistency. Alven completed Task 1, and Ben completed Task 2.

## Front Matter

```{r}
rm(list = ls())
library(tidyverse)
library(readxl)
library(kableExtra)
library(forcats)

cod <- read_excel("CODGames2_mp.xlsx", 
  col_types = c("text", "text", "text", 
  "text", "date", "text", "text", "numeric", 
  "numeric", "numeric", "numeric", 
  "numeric", "text", "text", "text", 
  "text", "numeric", "numeric", "numeric", 
  "numeric", "numeric", "numeric", 
  "numeric", "numeric", "numeric", 
  "numeric", "numeric"))

```

## Task 1 - Analysis of TotalXP with XPTYpes

### Part 1 - Remove Matches with Partial Participation

### Part 2 - Side-by-Side Boxplot Analysis of TotalXP and XPType

### Part 3 - Supplementary Summary Statistics

### Part 4 - Relationship between XPType and TotalXP

## Task 2 - Model for Score Variable in Team Death Match

Of the predictors total XP, eliminations, deaths, damage, XP type, and whether
the player's team won, which should be included in the model?

### Part 1 - LASSO Regression and 1 other Feature Selection Procedure (From Lecture 15)

Include relevant plots, discussion of lambda, estimated equation from LASSO and 
other method. Discuss/compare results.

### Part 2 - Regression Tree for Predicting Score

Include total XP, eliminations, deaths, damage, XPType,  and whether the
player's team won. Each node must contain >=15 obs.

Display the tree and report the variables associated with the 3 highest variable
importance values.

### Part 3 - Standardized Input Regression Model

Based on the variables from the "other method" in Task 2 - Part 1: standardize 
the inputs, build the regression model, report the estimated equation, and 
report the 3 most important variables based on the magnitude of the estimated
coefficients.

Compare the 3 most important variables with the regression tree's 3 highest 
variable importance values.
